Title

### 2012-01-20 inc/laser/laser.tex
Began porting OE references into laser.tex. Currently through first 5

% page 8630
Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
As with other Yb-doped laser crystals [1], this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load. 
Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated [2-5] and are now available commercially [6,7].
Such lasers have used dichroic mirrors [2,4,8,9], polarization-coupling [5], or more complex optical geometries (e.g., thin disk lasers [6,10]) to pump the laser crystal in an efficient manner.

% page 8631
The spectroscopic properties of the biaxial Yb:KGW laser crystal [11] indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength [1].
However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) [12,13], mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis [5,11]. This optimal requirement for orthogonal linear pump and laser polarizations led Holtom [5] to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.

@@

The design of the laser head follows the example of Holtom [5], which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism [16].
To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
On the other hand, as with \ce{Yb$:$KY(WO4)2} [12,13], the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis [5]) in this biaxial material. 
%page 8632
Instead of using a polarization-coupled pumping geometry [5] to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig. 1(a), where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (Fig. 1(b)); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 [11,15].
The \ce{SiO2} coating also has an appropriate bandwidth for femtosecond laser operation in excess of 100nm (to the
0.1\% reflection points) for $p$-polarized radiation centered at 1040nm.
On the other hand, the $\sim$ 0.4\% reflection loss per surface for $s$-polarized 1040nm radiation is sufficient to suppress $s$-polarized oscillation in a laser resonator.
In addition, the same anti-reflection coating has less than 0.5\% reflection loss for $s$-polarized 980nm pump radiation (polarized parallel to the $a$-axis of the Yb:KGW crystal) at incidence angles between 21 and 39$^\circ$ (Fig. 1(b)).
Thus, in this Yb:KGW laser crystal geometry, the largest absorption cross-section associated with the optical $N_m$-axis is accessed with about 94\% efficiency ($\cos 20^\circ = 0.94$) by the $s$-polarized pump radiation.

@@

The spatial pump distribution needed to generate the required $B_T/A_T$ ratio can be found using a cosine Fourier series solution to the thermal diffusion equation for our two-dimensional problem,
\begin{equation} \label{eq:heat_eqn}
  \kappa \nabla^2 T(x,y) = Q(x,y)
  \quad\text{,}
\end{equation}
where $\kappa$ is the thermal diffusion coefficient.
Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry [16].
The specifications of our dual-axis collimated diodes [26] indicate that heat source can be written as
\begin{equation}
  Q(x,y) = Q_0 \exp \left[ - \frac{x^2}{W_x^2} - \left( \frac{y^2}{W_y^2} \right)^3 \right]
  \quad\text{;}
\end{equation}
that is, the focused pump radiation is described accurately by a Gaussian irradiance distribution in the horizontal ($x$) direction with half-width 1/e maximum (HW1/eM) spot size $W_x$ and a super-Gaussian of order 3 in the vertical ($y$) direction with spot size $W_y$.
\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique [16]; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
This analytical result for the central region around the laser beam axis is shown in Fig. 2; displayed are the
%page 8634
horizontal ($y = 0$, Fig. 2(a)) and vertical ($x = 0$, Fig. 2(b)) sections, respectively, through the $Q(x,y)$ and $T(x,y)$ distributions.
Also shown are the parabolic approximations to the pump-induced temperature distribution (dashed lines) and representative TEM$_{00}$ Gaussian laser modes (shaded area) in the Yb:KGW crystal with a HW1/eM field mode sizes of $w_y = 160\mu$m and $w_x = 182 \mu$m (14\% larger due to refraction (Fig. 1(a)).
This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium [1]; in particular, in the vertical direction (Fig. 2(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.

@@

The analysis clearly shows that the TEM$_{00}$ laser mode will experience a near perfect pump-induced thermal lens upon propagation through the gain medium --- the deviation of $T(x,y)$ from the parabolic approximation, and hence a perfect dominant thermo-optic (GRIN) duct, being insignificant over the size of the laser mode in both the horizontal and vertical directions.
Moreover, the fact that $W_x \approx 1.6 w_x$ implies that some degree of horizontal spatial walk-off between the pump and laser beams can be tolerated in this TLS geometry.
For example, when the pump and laser beams are exactly overlapped at the center of the Yb:KGW crystal, an external angle of 0.1rad (= 5.7$^\circ$) would generate an effective spatial walk-off of only about 75$\mu$m over our 3mm gain crystal length.
Clearly, this can be readily accommodated by the $W_x = 0.3$mm horizontal pump spot size when $w_x \approx 180\mu$m.
However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium [1].

We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) [11,14] and stress-induced refractive index changes [31].
For the employed crystal cut in our TLS geometry (Fig. 1(a)), the anisotropy in $\kappa$ [14,27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
The refractive index
%page 8635
changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index [21].
Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping [31,34].
Consequently, for our laser polarized along the optical $N_p$-axis [5], there is insufficient information to include accurately the effects of stress-induced refractive index changes.

@@

The stability analysis of our diode-pumped Yb:KGW laser (Fig. 3(b)), clearly indicates that the resonator configuration is unstable for thermal focal lengths less than about $-6m^{-1}$ (diopters) and is unlikely to oscillate for negative thermal focal lengths as this requires a vertical TEM$_{00}$-mode size $w_y$ greater than 180$\mu$m in the Yb:KGW gain medium --- a value that would result in strong absorptive losses due to the restrictive $W_y = 200\mu$m super-Gaussian vertical pump beam size in the Yb-doped crystal.
The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW [11], but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis [5,34].

@@

In the cavity configuration depicted in Fig. 3(a), self-starting mode-locked (ML) operation was observed for output powers greater than $\sim$2W and $z \approx 42$cm.
The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (Fig. 4(a)).
The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH [32]) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (Fig. 4(b)) at an optimum mode-locked output power of $5.0$W.
This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity [2].

>>>>

% page 8630
Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load. 
Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially [6,7].
Such lasers have used dichroic mirrors\cite{Brunner_diode_pumped,Major_femtosecond_2006} [8,9], polarization-coupling\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers [6,10]) to pump the laser crystal in an efficient manner.

% page 8631
The spectroscopic properties of the biaxial Yb:KGW laser crystal [11] indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength\cite{Brenier_new_criteria}.
However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) [12,13], mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis\cite{Holtom_mode_locked_2006} [11]. This optimal requirement for orthogonal linear pump and laser polarizations led Holtom\cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.

@@

The design of the laser head follows the example of Holtom\cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism [16].
To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
On the other hand, as with \ce{Yb$:$KY(WO4)2} [12,13], the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis\cite{Holtom_mode_locked_2006}) in this biaxial material. 
%page 8632
Instead of using a polarization-coupled pumping geometry\cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig. 1(a), where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (Fig. 1(b)); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 [11,15].
The \ce{SiO2} coating also has an appropriate bandwidth for femtosecond laser operation in excess of 100nm (to the
0.1\% reflection points) for $p$-polarized radiation centered at 1040nm.
On the other hand, the $\sim$ 0.4\% reflection loss per surface for $s$-polarized 1040nm radiation is sufficient to suppress $s$-polarized oscillation in a laser resonator.
In addition, the same anti-reflection coating has less than 0.5\% reflection loss for $s$-polarized 980nm pump radiation (polarized parallel to the $a$-axis of the Yb:KGW crystal) at incidence angles between 21 and 39$^\circ$ (Fig. 1(b)).
Thus, in this Yb:KGW laser crystal geometry, the largest absorption cross-section associated with the optical $N_m$-axis is accessed with about 94\% efficiency ($\cos 20^\circ = 0.94$) by the $s$-polarized pump radiation.

@@

The spatial pump distribution needed to generate the required $B_T/A_T$ ratio can be found using a cosine Fourier series solution to the thermal diffusion equation for our two-dimensional problem,
\begin{equation} \label{eq:heat_eqn}
  \kappa \nabla^2 T(x,y) = Q(x,y)
  \quad\text{,}
\end{equation}
where $\kappa$ is the thermal diffusion coefficient.
Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry [16].
The specifications of our dual-axis collimated diodes [26] indicate that heat source can be written as
\begin{equation}
  Q(x,y) = Q_0 \exp \left[ - \frac{x^2}{W_x^2} - \left( \frac{y^2}{W_y^2} \right)^3 \right]
  \quad\text{;}
\end{equation}
that is, the focused pump radiation is described accurately by a Gaussian irradiance distribution in the horizontal ($x$) direction with half-width 1/e maximum (HW1/eM) spot size $W_x$ and a super-Gaussian of order 3 in the vertical ($y$) direction with spot size $W_y$.
\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique [16]; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
This analytical result for the central region around the laser beam axis is shown in Fig. 2; displayed are the
%page 8634
horizontal ($y = 0$, Fig. 2(a)) and vertical ($x = 0$, Fig. 2(b)) sections, respectively, through the $Q(x,y)$ and $T(x,y)$ distributions.
Also shown are the parabolic approximations to the pump-induced temperature distribution (dashed lines) and representative TEM$_{00}$ Gaussian laser modes (shaded area) in the Yb:KGW crystal with a HW1/eM field mode sizes of $w_y = 160\mu$m and $w_x = 182 \mu$m (14\% larger due to refraction (Fig. 1(a)).
This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium\cite{Brenier_new_criteria}; in particular, in the vertical direction (Fig. 2(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.

@@

The analysis clearly shows that the TEM$_{00}$ laser mode will experience a near perfect pump-induced thermal lens upon propagation through the gain medium --- the deviation of $T(x,y)$ from the parabolic approximation, and hence a perfect dominant thermo-optic (GRIN) duct, being insignificant over the size of the laser mode in both the horizontal and vertical directions.
Moreover, the fact that $W_x \approx 1.6 w_x$ implies that some degree of horizontal spatial walk-off between the pump and laser beams can be tolerated in this TLS geometry.
For example, when the pump and laser beams are exactly overlapped at the center of the Yb:KGW crystal, an external angle of 0.1rad (= 5.7$^\circ$) would generate an effective spatial walk-off of only about 75$\mu$m over our 3mm gain crystal length.
Clearly, this can be readily accommodated by the $W_x = 0.3$mm horizontal pump spot size when $w_x \approx 180\mu$m.
However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium\cite{Brenier_new_criteria}.

We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) [11,14] and stress-induced refractive index changes [31].
For the employed crystal cut in our TLS geometry (Fig. 1(a)), the anisotropy in $\kappa$ [14,27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
The refractive index
%page 8635
changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index [21].
Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping [31,34].
Consequently, for our laser polarized along the optical $N_p$-axis\cite{Holtom_mode_locked_2006}, there is insufficient information to include accurately the effects of stress-induced refractive index changes.

@@

The stability analysis of our diode-pumped Yb:KGW laser (Fig. 3(b)), clearly indicates that the resonator configuration is unstable for thermal focal lengths less than about $-6m^{-1}$ (diopters) and is unlikely to oscillate for negative thermal focal lengths as this requires a vertical TEM$_{00}$-mode size $w_y$ greater than 180$\mu$m in the Yb:KGW gain medium --- a value that would result in strong absorptive losses due to the restrictive $W_y = 200\mu$m super-Gaussian vertical pump beam size in the Yb-doped crystal.
The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW [11], but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis\cite{Holtom_mode_locked_2006} [34].

@@

In the cavity configuration depicted in Fig. 3(a), self-starting mode-locked (ML) operation was observed for output powers greater than $\sim$2W and $z \approx 42$cm.
The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (Fig. 4(a)).
The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH [32]) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (Fig. 4(b)) at an optimum mode-locked output power of $5.0$W.
This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity\cite{Brunner_diode_pumped}.
